---
title: "NLPShinyApp"
author: "KKher"
date: "9/18/2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, error=FALSE, warning=FALSE)
```

## Data

our data comes from [HC Corpora](http://www.corpora.heliohost.org), and can be downloaded directly from [here](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip)

First, we will read all our files:
```{r gettingData}
US_blogs <- readLines("./final/en_US/en_US.blogs.txt")
US_news <- readLines("./final/en_US/en_US.news.txt")
US_twitter <- readLines("./final/en_US/en_US.twitter.txt")
```

## Objectives

1. A Shiny app that takes as input a phrase (multiple words) in a text box input and outputs a prediction of the next word.
2. A slide deck consisting of no more than 5 slides created with [R Studio Presenter](https://support.rstudio.com/hc/en-us/articles/200486468-Authoring-R-Presentations) pitching your algorithm and app as if you were presenting to your boss or an investor.

## Data Sampling

Since working on the entire dataset will require high computational power & will consume time, -due to their sizes-, we will work on on samples of our data, exactly 10% of each document size.

```{r dataSamples}
library(quanteda)

set.seed(18092020)
# will take samples per country
sampleUS_blogs <- sample(US_blogs,length(US_blogs) * 0.1, replace = FALSE)
sampleUS_news <- sample(US_news,length(US_news) * 0.1, replace = FALSE)
sampleUS_twitter <- sample(US_twitter,length(US_twitter) * 0.1, replace = FALSE)

# combine samples into one ENGLISH sample set
en_SampledData <- paste(sampleUS_blogs, sampleUS_news, sampleUS_twitter)

# create corpus for the combined and individual samples
us_corp <- corpus(en_SampledData)
# blogs_corp <- corpus(sampleUS_blogs);news_corp <- corpus(sampleUS_news);twitter_corp <- corpus(sampleUS_twitter)

# remove unneeded variables
remove(sampleUS_blogs, sampleUS_news, sampleUS_twitter,en_SampledData,
       US_blogs, US_news, US_twitter)
```

## Data Preprocessing

After we've our corpuses set, we need to clean them from any non-alphabetic entries

```{r dataCleaning}
# the puncuations and numbers in the texts were removed as there is no need to predict punctations or numbers
us_corp <- tokens(
    x = tolower(us_corp),
    remove_punct = TRUE,
    remove_numbers = TRUE,
    remove_symbols = TRUE,
    remove_url = TRUE
)

# blogs_corp <- tokens(
#     x = tolower(blogs_corp),
#     remove_punct = TRUE,
#     remove_numbers = TRUE,
#     remove_symbols = TRUE,
#     remove_url = TRUE
# )
# 
# news_corp <- tokens(
#     x = tolower(news_corp),
#     remove_punct = TRUE,
#     remove_numbers = TRUE,
#     remove_symbols = TRUE,
#     remove_url = TRUE
# )
# 
# twitter_corp <- tokens(
#     x = tolower(twitter_corp),
#     remove_punct = TRUE,
#     remove_numbers = TRUE,
#     remove_symbols = TRUE,
#     remove_url = TRUE
# )

```

## Data Tokenization and N-Grams

Generating 2, 3, and 4-grams for combined and individual sets

```{r corpusNGrams}
us2gram <- tokens_ngrams(us_corp, n = 2)
us3gram <- tokens_ngrams(us_corp, n = 3)
us4gram <- tokens_ngrams(us_corp, n = 4)
# 
# blogs2gram <- tokens_ngrams(blogs_corp, n = 2)
# blogs3gram <- tokens_ngrams(blogs_corp, n = 3)
# 
# news2gram <- tokens_ngrams(news_corp, n = 2)
# news3gram <- tokens_ngrams(news_corp, n = 3)
# 
# twitter2gram <- tokens_ngrams(twitter_corp, n = 2)
# twitter3gram <- tokens_ngrams(twitter_corp, n = 3)

```

Create Document Frequency Matrices and trim them to speedup the process

```{r dataDFM}
# All data DFMs
us_uni_DFM <- dfm(us_corp) ; us_uni_DFM <- dfm_trim(us_uni_DFM, 10)
us_bi_DFM <- dfm(us2gram) ; us_bi_DFM <- dfm_trim(us_bi_DFM, 10)
us_tri_DFM <- dfm(us3gram) ; us_tri_DFM <- dfm_trim(us_tri_DFM, 10)
us_quad_DFM <- dfm(us4gram) ; us_quad_DFM <- dfm_trim(us_quad_DFM, 10)

# # News data DFMs
# news_uni_DFM <- dfm(news_corp) ; news_uni_DFM <- dfm_trim(news_uni_DFM, 3)
# news_bi_DFM <- dfm(news2gram) ; news_bi_DFM <- dfm_trim(news_bi_DFM, 3)
# news_tri_DFM <- dfm(news3gram) ; news_tri_DFM <- dfm_trim(news_tri_DFM, 3)
# 
# # Blogs data DFMs
# blogs_uni_DFM <- dfm(blogs_corp) ; blogs_uni_DFM <- dfm_trim(blogs_uni_DFM, 3)
# blogs_bi_DFM <- dfm(blogs2gram) ; blogs_bi_DFM <- dfm_trim(blogs_bi_DFM, 3)
# blogs_tri_DFM <- dfm(blogs3gram) ; blogs_tri_DFM <- dfm_trim(blogs_tri_DFM, 3)
# 
# # Twitter data DFMs
# twitt_uni_DFM <- dfm(twitter_corp) ; twitt_uni_DFM <- dfm_trim(twitt_uni_DFM, 3)
# twitt_bi_DFM <- dfm(twitter2gram) ; twitt_bi_DFM <- dfm_trim(twitt_bi_DFM, 3)
# twitt_tri_DFM <- dfm(twitter3gram) ; twitt_tri_DFM <- dfm_trim(twitt_tri_DFM, 3)

# remove unneeded variables
remove(us_corp, us2gram, us3gram,us4gram)

# remove(us_corp, us2gram, us3gram,news_corp,news2gram,news3gram,blogs_corp,blogs2gram,blogs3gram,
#        twitter_corp, twitter2gram, twitter3gram)
```

Convert DFMs into tables with words and their respective counts

```{r wordFrequency_us}
library(data.table)

us_uni_sum <- colSums(us_uni_DFM)
us_bi_sum <- colSums(us_bi_DFM)
us_tri_sum <- colSums(us_tri_DFM)
us_quad_sum <- colSums(us_quad_DFM)

# Create data tables with individual words as columns
us_uni_words <- data.table(word_1 = names(us_uni_sum), count = us_uni_sum)

us_bi_words <- data.table(
        word_1 = sapply(strsplit(names(us_bi_sum), "_", fixed = TRUE), '[[', 1),
        word_2 = sapply(strsplit(names(us_bi_sum), "_", fixed = TRUE), '[[', 2),
        count = us_bi_sum)

us_tri_words <- data.table(
        word_1 = sapply(strsplit(names(us_tri_sum), "_", fixed = TRUE), '[[', 1),
        word_2 = sapply(strsplit(names(us_tri_sum), "_", fixed = TRUE), '[[', 2),
        word_3 = sapply(strsplit(names(us_tri_sum), "_", fixed = TRUE), '[[', 3),
        count = us_tri_sum)

us_quad_words <- data.table(
        word_1 = sapply(strsplit(names(us_quad_sum), "_", fixed = TRUE), '[[', 1),
        word_2 = sapply(strsplit(names(us_quad_sum), "_", fixed = TRUE), '[[', 2),
        word_3 = sapply(strsplit(names(us_quad_sum), "_", fixed = TRUE), '[[', 3),
        word_4 = sapply(strsplit(names(us_quad_sum), "_", fixed = TRUE), '[[', 4),
        count = us_tri_sum)

```

```{r wordFrequency_blogs}
# blogs_uni_sum <- colSums(blogs_uni_DFM)
# blogs_bi_sum <- colSums(blogs_bi_DFM)
# blogs_tri_sum <- colSums(blogs_tri_DFM)
# blogs_quad_sum <- colSums(blogs_quad_DFM)
# 
# # Create data tables with individual words as columns
# blogs_uni_words <- data.table(word_1 = names(blogs_uni_sum), count = blogs_uni_sum)
# 
# blogs_bi_words <- data.table(
#         word_1 = sapply(strsplit(names(blogs_bi_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(blogs_bi_sum), "_", fixed = TRUE), '[[', 2),
#         count = blogs_bi_sum)
# 
# blogs_tri_words <- data.table(
#         word_1 = sapply(strsplit(names(blogs_tri_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(blogs_tri_sum), "_", fixed = TRUE), '[[', 2),
#         word_3 = sapply(strsplit(names(blogs_tri_sum), "_", fixed = TRUE), '[[', 3),
#         count = blogs_tri_sum)
# 

```

```{r wordFrequency_news}
# news_uni_sum <- colSums(news_uni_DFM)
# news_bi_sum <- colSums(news_bi_DFM)
# news_tri_sum <- colSums(news_tri_DFM)
# 
# # Create data tables with individual words as columns
# news_uni_words <- data.table(word_1 = names(news_uni_sum), count = news_uni_sum)
# 
# news_bi_words <- data.table(
#         word_1 = sapply(strsplit(names(news_bi_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(news_bi_sum), "_", fixed = TRUE), '[[', 2),
#         count = news_bi_sum)
# 
# news_tri_words <- data.table(
#         word_1 = sapply(strsplit(names(news_tri_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(news_tri_sum), "_", fixed = TRUE), '[[', 2),
#         word_3 = sapply(strsplit(names(news_tri_sum), "_", fixed = TRUE), '[[', 3),
#         count = news_tri_sum)


```

```{r wordFrequency_twitter}
# twitt_uni_sum <- colSums(twitt_uni_DFM)
# twitt_bi_sum <- colSums(twitt_bi_DFM)
# twitt_tri_sum <- colSums(twitt_tri_DFM)
# 
# # Create data tables with individual words as columns
# twitt_uni_words <- data.table(word_1 = names(twitt_uni_sum), count = twitt_uni_sum)
# 
# twitt_bi_words <- data.table(
#         word_1 = sapply(strsplit(names(twitt_bi_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(twitt_bi_sum), "_", fixed = TRUE), '[[', 2),
#         count = twitt_bi_sum)
# 
# twitt_tri_words <- data.table(
#         word_1 = sapply(strsplit(names(twitt_tri_sum), "_", fixed = TRUE), '[[', 1),
#         word_2 = sapply(strsplit(names(twitt_tri_sum), "_", fixed = TRUE), '[[', 2),
#         word_3 = sapply(strsplit(names(twitt_tri_sum), "_", fixed = TRUE), '[[', 3),
#         count = twitt_tri_sum)


```

## Data Modelling
Let’s add Kneser-Kney smoothing to the dataset. First we will find bi-gram probabilities and then add smoothing.

```{r dataModelling_bi}
# remove unneeded variables
remove(us_uni_sum,us_bi_sum, us_tri_sum, us_quad_sum, us_uni_DFM,us_bi_DFM,us_tri_DFM,us_quad_DFM)

discount_value <- 0.75

######## Finding Bi-Gram Probability for Combined Data #################

# Finding number of bi-gram words
numOfBiGrams <- nrow(us_bi_words[by = .(word_1, word_2)])

# Dividing number of times word 2 occurs as second part of bigram, by total number of bigrams.  
# ( Finding probability for a word given the number of times it was second word of a bigram)
ckn <- us_bi_words[, .(Prob = ((.N) / numOfBiGrams)), by = word_2]
setkey(ckn, word_2)

# Assigning the probabilities as second word of bigram, to unigrams
us_uni_words[, Prob := ckn[word_1, Prob]]
us_uni_words <- us_uni_words[!is.na(us_uni_words$Prob)]

# Finding number of times word 1 occurred as word 1 of bi-grams
n1wi <- us_bi_words[, .(count = .N), by = word_1]
setkey(n1wi, word_1)
setkey(us_uni_words, word_1)

# Assigning total times word 1 occured to bigram cn1
us_bi_words[, Cn1 := n1wi[word_1, count]]

# Kneser Kney Algorithm
us_bi_words[, Prob := ((count - discount_value) / Cn1 + discount_value / Cn1 * n1wi[word_1, count] * us_uni_words[word_2, Prob])]

######## End of Finding Bi-Gram Probability for Combined Data #################

######## Finding Bi-Gram Probability for News Data #################

# # Finding number of bi-gram words
# numOfBiGrams <- nrow(news_bi_words[by = .(word_1, word_2)])
# 
# # Dividing number of times word 2 occurs as second part of bigram, by total number of bigrams.  
# # ( Finding probability for a word given the number of times it was second word of a bigram)
# ckn <- news_bi_words[, .(Prob = ((.N) / numOfBiGrams)), by = word_2]
# setkey(ckn, word_2)
# 
# # Assigning the probabilities as second word of bigram, to unigrams
# news_uni_words[, Prob := ckn[word_1, Prob]]
# news_uni_words <- news_uni_words[!is.na(news_uni_words$Prob)]
# 
# # Finding number of times word 1 occurred as word 1 of bi-grams
# n1wi <- news_bi_words[, .(count = .N), by = word_1]
# setkey(n1wi, word_1)
# setkey(news_uni_words, word_1)
# 
# # Assigning total times word 1 occured to bigram cn1
# news_bi_words[, Cn1 := n1wi[word_1, count]]
# 
# # Kneser Kney Algorithm
# news_bi_words[, Prob := ((count - discount_value) / Cn1 + discount_value / Cn1 * n1wi[word_1, count] * news_uni_words[word_2, Prob])]

######## End of Finding Bi-Gram Probability for News Data #################

######## Finding Bi-Gram Probability for Blogs Data #################

# Finding number of bi-gram words
# numOfBiGrams <- nrow(blogs_bi_words[by = .(word_1, word_2)])
# 
# # Dividing number of times word 2 occurs as second part of bigram, by total number of bigrams.  
# # ( Finding probability for a word given the number of times it was second word of a bigram)
# ckn <- blogs_bi_words[, .(Prob = ((.N) / numOfBiGrams)), by = word_2]
# setkey(ckn, word_2)
# 
# # Assigning the probabilities as second word of bigram, to unigrams
# blogs_uni_words[, Prob := ckn[word_1, Prob]]
# blogs_uni_words <- blogs_uni_words[!is.na(blogs_uni_words$Prob)]
# 
# # Finding number of times word 1 occurred as word 1 of bi-grams
# n1wi <- blogs_bi_words[, .(count = .N), by = word_1]
# setkey(n1wi, word_1)
# setkey(blogs_uni_words, word_1)
# 
# # Assigning total times word 1 occured to bigram cn1
# blogs_bi_words[, Cn1 := n1wi[word_1, count]]
# 
# # Kneser Kney Algorithm
# blogs_bi_words[, Prob := ((count - discount_value) / Cn1 + discount_value / Cn1 * n1wi[word_1, count] * blogs_uni_words[word_2, Prob])]

######## End of Finding Bi-Gram Probability for Blogs Data #################


######## Finding Bi-Gram Probability for Twitter Data #################

# Finding number of bi-gram words
# numOfBiGrams <- nrow(twitt_bi_words[by = .(word_1, word_2)])
# 
# # Dividing number of times word 2 occurs as second part of bigram, by total number of bigrams.  
# # ( Finding probability for a word given the number of times it was second word of a bigram)
# ckn <- twitt_bi_words[, .(Prob = ((.N) / numOfBiGrams)), by = word_2]
# setkey(ckn, word_2)
# 
# # Assigning the probabilities as second word of bigram, to unigrams
# twitt_uni_words[, Prob := ckn[word_1, Prob]]
# twitt_uni_words <- twitt_uni_words[!is.na(twitt_uni_words$Prob)]
# 
# # Finding number of times word 1 occurred as word 1 of bi-grams
# n1wi <- twitt_bi_words[, .(count = .N), by = word_1]
# setkey(n1wi, word_1)
# setkey(twitt_uni_words, word_1)
# 
# # Assigning total times word 1 occured to bigram cn1
# twitt_bi_words[, Cn1 := n1wi[word_1, count]]
# 
# # Kneser Kney Algorithm
# twitt_bi_words[, Prob := ((count - discount_value) / Cn1 + discount_value / Cn1 * n1wi[word_1, count] * twitt_uni_words[word_2, Prob])]

######## End of Finding Bi-Gram Probability for Twitter Data #################
```

Then let’s find tri-gram probabilities and add smoothing.

```{r dataModelling_tri}
######## Finding Tri-Gram Probability for Combined Data #################

# Finding count of word1-word2 combination in bigram 
setDT(us_tri_words)[us_bi_words, Cn2 := count, on = .(word_1 = word_1, word_2 = word_2)]
setDT(us_tri_words)[us_bi_words, Prob := Prob, on = .(word_1 = word_1, word_2 = word_2)]

#us_tri_words[, Cn2 := us_bi_words[.(word_1, word_2), count]]

# Finding count of word1-word2 combination in trigram
n1w12 <- us_tri_words[, .N, by = .(word_1, word_2)]
setkey(n1w12, word_1, word_2)

setDT(us_tri_words)[n1w12, N := N, on = .(word_1 = word_1, word_2 = word_2)]

# Kneser Kney Algorithm
us_tri_words[, Prob := (count - discount_value) / Cn2 + discount_value / Cn2 * N* Prob]

######## End of Finding Tri-Gram Probability for Combined Data #################

###################################################################################################
######## Finding Tri-Gram Probability for News Data #################

# Finding count of word1-word2 combination in bigram 
# setDT(news_tri_words)[news_bi_words, Cn2 := count, on = .(word_1 = word_1, word_2 = word_2)]
# setDT(news_tri_words)[news_bi_words, Prob := Prob, on = .(word_1 = word_1, word_2 = word_2)]
# 
# #us_tri_words[, Cn2 := us_bi_words[.(word_1, word_2), count]]
# 
# # Finding count of word1-word2 combination in trigram
# n1w12 <- news_tri_words[, .N, by = .(word_1, word_2)]
# setkey(news_bi_words, word_1, word_2)
# 
# setDT(news_tri_words)[n1w12, N := N, on = .(word_1 = word_1, word_2 = word_2)]
# 
# # Kneser Kney Algorithm
# news_tri_words[, Prob := (count - discount_value) / Cn2 + discount_value / Cn2 * N * Prob]

######## End of Finding Tri-Gram Probability for News Data #################

###################################################################################################
######## Finding Tri-Gram Probability for Blogs Data #################

# Finding count of word1-word2 combination in bigram 
# setDT(blogs_tri_words)[blogs_bi_words, Cn2 := count, on = .(word_1 = word_1, word_2 = word_2)]
# setDT(blogs_tri_words)[blogs_bi_words, Prob := Prob, on = .(word_1 = word_1, word_2 = word_2)]
# 
# #us_tri_words[, Cn2 := us_bi_words[.(word_1, word_2), count]]
# 
# # Finding count of word1-word2 combination in trigram
# n1w12 <- blogs_tri_words[, .N, by = .(word_1, word_2)]
# setkey(n1w12, word_1, word_2)
# 
# setDT(blogs_tri_words)[n1w12, N := N, on = .(word_1 = word_1, word_2 = word_2)]
# 
# # Kneser Kney Algorithm
# blogs_tri_words[, Prob := (count - discount_value) / Cn2 + discount_value / Cn2 * N * Prob]

######## End of Finding Tri-Gram Probability for Blogs Data #################

###################################################################################################
######## Finding Tri-Gram Probability for Twitter Data #################

# Finding count of word1-word2 combination in bigram 
# setDT(twitt_tri_words)[twitt_bi_words, Cn2 := count, on = .(word_1 = word_1, word_2 = word_2)]
# setDT(twitt_tri_words)[twitt_bi_words, Prob := Prob, on = .(word_1 = word_1, word_2 = word_2)]
# 
# #us_tri_words[, Cn2 := us_bi_words[.(word_1, word_2), count]]
# 
# # Finding count of word1-word2 combination in trigram
# n1w12 <- twitt_tri_words[, .N, by = .(word_1, word_2)]
# setkey(n1w12, word_1, word_2)
# 
# setDT(twitt_tri_words)[n1w12, N := N, on = .(word_1 = word_1, word_2 = word_2)]
# 
# # Kneser Kney Algorithm
# twitt_tri_words[, Prob := (count - discount_value) / Cn2 + discount_value / Cn2 * N * Prob]

######## End of Finding Tri-Gram Probability for Twitter Data #################


```

```{r dataModelling_quad}
######## Finding Tri-Gram Probability for Combined Data #################

# Finding count of word1-word2 combination in bigram 
setDT(us_quad_words)[us_tri_words, Cn2 := count, on = .(word_1 = word_1, word_2 = word_2, word_3 = word_3)]
setDT(us_quad_words)[us_tri_words, Prob := Prob, on = .(word_1 = word_1, word_2 = word_2, word_3 = word_3)]

#us_tri_words[, Cn2 := us_bi_words[.(word_1, word_2), count]]

# Finding count of word1-word2 combination in trigram
n1w12 <- us_quad_words[, .N, by = .(word_1, word_2, word_3)]
setkey(n1w12, word_1, word_2, word_3)

setDT(us_quad_words)[n1w12, N := N, on = .(word_1 = word_1, word_2 = word_2, word_3 = word_3)]

# Kneser Kney Algorithm
us_quad_words[, Prob := (count - discount_value) / Cn2 + discount_value / Cn2 * N* Prob]

######## End of Finding Tri-Gram Probability for Combined Data #################

```

## Saving results

We now save 1,2,3, 4-grams outputs, to furtherly utilize them in our ShinyApp
```{r dataSaving}
saveRDS(us_uni_words, "./NLPData/us_uni_words.RData")
saveRDS(us_bi_words, "./NLPData/us_bi_words.RData")
saveRDS(us_tri_words, "./NLPData/us_tri_words.RData")
saveRDS(us_quad_words, "./NLPData/us_quad_words.RData")

# saveRDS(news_uni_words, "./NLPData/news_uni_words.RData")
# saveRDS(news_bi_words, "./NLPData/news_bi_words.RData")
# saveRDS(news_tri_words, "./NLPData/news_tri_words.RData")
# 
# saveRDS(blogs_uni_words, "./NLPData/blogs_uni_words.RData")
# saveRDS(blogs_bi_words, "./NLPData/blogs_bi_words.RData")
# saveRDS(blogs_tri_words, "./NLPData/blogs_tri_words.RData")
# 
# saveRDS(twitt_uni_words, "./NLPData/twitt_uni_words.RData")
# saveRDS(twitt_bi_words, "./NLPData/twitt_bi_words.RData")
# saveRDS(twitt_tri_words, "./NLPData/twitt_tri_words.RData")

```
